---
title: "Computing plasticity indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing plasticity indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette walks through computing a handful of `ppindices` plasticity
indices on a small, toy reaction norm. A reaction norm here is simply a set of
trait measurements for one genotype, taken across several environments.

```{r setup}
library(ppindices)
```

## A toy reaction norm

Suppose a genotype was measured in four environments, with trait values
increasing steadily as the environment changes:

```{r}
trait_values <- c(2, 4, 6, 8)
environments <- c(1, 2, 3, 4)
```

## Coefficient of variation of trait values

`calculate_CVt()` computes the coefficient of variation of the trait values
across environments — a simple, scale-independent measure of overall
variability:

```{r}
calculate_CVt(trait_values)
```

## Reaction norm slope

`calculate_reaction_norm_slope()` fits a linear regression of the trait
values on the environmental values and returns the slope, describing the
linear rate of phenotypic change across environments:

```{r}
calculate_reaction_norm_slope(trait_values, environments = environments)
```

## Relative distance plasticity index (RDPI)

`calculate_rdpi()` averages the relative distance between trait values over
all pairs of environments:

```{r}
calculate_rdpi(trait_values, env_values = environments)
```

## Environment-specific plasticity (ESP)

`calculate_ESP()` sums the absolute relative deviations of each
environment's mean trait value from the overall mean. Here each environment
label is repeated to represent two replicate measurements per environment:

```{r}
env_labels <- c(1, 1, 2, 2)
calculate_ESP(trait_values, env_values = env_labels)
```

## Summary

These four indices illustrate the range of approaches implemented in
`ppindices`: variance-based (`calculate_CVt`), slope-based
(`calculate_reaction_norm_slope`), distance-based (`calculate_rdpi`), and
index-based (`calculate_ESP`) measures of phenotypic plasticity. The package
provides 43 such indices in total; see `help(package = "ppindices")` for the
complete list.
